### 2014-10-31 diary/10_Diary_1.tex
adds diary structure & raw report about the first two sessions

\documentclass[a4paper,11pt]{article}

\usepackage{graphicx}
\usepackage{natbib}
\usepackage[utf8]{inputenc}
\usepackage{tabularx}
\usepackage{hyperref}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
% \usepackage{mathptmx} % Times New Roman

\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\footskip}{15mm}

\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}

% --------------------------------------------------------------

\providecommand*{\ShortTitle}{WaveMeIn}
\providecommand*{\FullTitle}{Projektprotokoll}
\providecommand*{\includeWiki}[1]{\wikiEnvironments \include{#1}\nowikiEnvironments}

% --------------------------------------------------------------

\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
{\textbf{\sffamily\Large \FullTitle}}
\vspace{1cm}}

\author{
{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
Group: 10\bigskip \\
Belk Stefan \\ {\small 0750926, 937, \href{mailto:belk.stefan@gmail.com}{belk.stefan@gmail.com}}\\
Petz Thomas \\ {\small 0601280, 937, \href{mailto:e0601280@student.tuwien.ac.at}{e0601280@student.tuwien.ac.at}}\\
Causevic Alma \\ {\small 0847805, 534, \href{mailto:alma.causevic@hotmail.com}{alma.causevic@hotmail.com}}\\ 
Causevic Amra  \\ {\small 0649241, 534, \href{mailto:amra.causevic@hotmail.com}{amra.causevic@hotmail.com}}\\ 
Seebacher David \\ {\small 0327243, 534, \href{mailto:david.seebacher@student.tuwien.ac.at}{david.seebacher@student.tuwien.ac.at}}\\
\vspace{4cm}
}

\begin{document}

\begin{titlepage}
\maketitle

\end{titlepage}

% --------------------------------------------------------------

%\input{diary/28.10.md}
%\input{diary/31.10.md}
\input{diary_bundle.tex}
\end{document}