### 2014-12-09 proposal/sections/workplan.tex
add workplan from googledoc to tex.

\section{Detailed description of the workpackages}
\label{sect:workplan}

\note{
\begin{itemize}
  \item {\em Length: 2-4 pages}
  \item Structuring the project into self-contained parts.
  \item Additional verbal descriptions.
  \item Work packages
  \begin{itemize}
    \item title
    \item goal(s)
    \item description
    \item expected results
    \item responsible person(s)
    \item dependencies
  \end{itemize}
\end{itemize}
}

\subsection{Basic research}
\begin{itemize}
 \item goal(s): The goal of the research part is to clarify the scientific requirements, needed for the creation of the prototype.
 \item description: This work package deals with the brain waves related part of this project.
 \item expected results: Thought identification, confirmation of the uniqueness of brain waves, measuring brain waves from different body parts.
 \item responsible person: Neuroscience researcher
 \item dependencies: none
 \item start date: 05.01.2015
 \item end date: 05.07.2016
\end{itemize}

\subsubsection{Confirm uniqueness of a person’s brain wave pattern}
\begin{itemize}
 \item goal(s): Confirm assumption of persons unique brain waves.
 \item description: Every person has an unique brain wave pattern for the same thought. This pattern can be used to identify the person.
 \item expected results: If the research shows unexpected results and it the uniqueness can’t be proved, it will limit the security aspect of the project, but it still can be realised.
 \item responsible person: Neuroscience researcher
 \item dependencies: none
 \item start date: 05.01.2015
 \item end date: 05.07.2015
\end{itemize}

\subsubsection{Reliable thought identification}
\begin{itemize}
 \item goal(s): Confirmation of our assumption, that a thought of a person can be identified again as the same.
 \item description: Research on how a specific brain wave of a person can be identified reliably.
 \item expected results: A person can be identified again as the same.
 \item responsible person: Neuroscience researcher
 \item dependencies: WP 1.1
 \item start date: 05.07.2015
 \item end date: 05.01.2016
\end{itemize}

\subsubsection{Measuring methods of brain waves from different parts of the head}
\begin{itemize}
 \item goal(s): Existing studies showed theoretical approaches to gatter brain waves with in-ear or behind-ear EEGs. The goal is to use this approaches and define requirements for building a small device to measure brain waves.
 \item description: In this package we attempt to define requirements for building a small device.
 \item expected results: Prove that brain waves can be measured from different head-parts than the forehead.
 \item responsible person: Neuroscience researcher
 \item dependencies: WP 1.2
 \item start date: 05.01.2016
 \item end date: 05.07.2016
\end{itemize}

\subsubsection{Survey of state of the art for sensors and devices for brain wave measurement}
\begin{itemize}
 \item goal(s): Analyse the hardware components of existing brain wave measurement and gather enough information to build a small device.
 \item description: There already exists a lot of sensors for detecting brain waves. Main part here is to look, if a device already exists or if we buy a existing one a make it smaller.
 \item expected results: Today do probably exist decent sensors, but none exactly as needed for the planned tiny device.
 \item responsible person: Hardware engineer
 \item dependencies: none
 \item start date: 05.01.2015
 \item end date: 05.01.2016
\end{itemize}

\subsection{Hardware and software development}
\begin{itemize}
 \item goal(s): Develop prototype hardware and software for showing the success.
 \item description: To show that authentication over brain waves works, we have to develop a hardware prototype. There different kinds of software we have to develop. First an algorithm to ensure fault tolerance and an interface for the communication with the device.
 \item expected results: The developed application shows a successfully communication with the device.
 \item responsible person: Hardware engineer
 \item dependencies: WP 1.2
 \item start date: 05.07.2016
 \item end date: 05.07.2017
\end{itemize}

\subsubsection{Reviewing of existing algorithms for detecting patterns in measured brain}
\begin{itemize}
 \item goal(s): To identify the thought of a person.
 \item description: It is possible that the same thought of a person can be shown different in the brain wave. For this case we need some patterns to confirm the measured brain waves.
 \item expected results: A working algorithm with good performance.
 \item responsible person: Computer science researcher
 \item dependencies: none
 \item start date: 05.07.2016
 \item end date: 05.10.2016
\end{itemize}

\subsubsection{Develop new algorithms for detecting patterns in measured brain waves}
\begin{itemize}
 \item goal(s):  Develop the needed algorithms.
 \item description: If there are no existing algorithms for detecting patterns in measured brain waves, which can be used in this project, then new ones need to be developed.
 \item expected results: Our algorithms need to be reliable and fast.
 \item responsible person: Computer science researcher
 \item dependencies: WP 2.1 
 \item start date: 05.10.2016
 \item end date: 05.01.2017
\end{itemize}

\subsubsection{Define the communication protocol and interface of the Wavy}
\begin{itemize}
 \item goal(s): A software interface for the Wavy
 \item description: Develop an software interface for other applications to communicate with the Wavy. This interface should operation system or other developers allow to handle their authentication over Wavy.
 \item expected results: A easy usable and reliable software interface implement for several programing languages.
 \item responsible person: Software engineer
 \item dependencies: none
 \item start date: 05.07.2016
 \item end date: 05.07.2017
\end{itemize}

\subsubsection{Develop small sensors for brain wave measurement}
\begin{itemize}
 \item goal(s): Develop small sensors for brain wave measurement so they can be placed in the Wavy.
 \item description: Because the Wavy should be build as tiny as possible small sensors are needed.
 \item expected results: This work results in a tiny fully functional sensor for surement.
 \item responsible person: Hardware engineer
 \item dependencies: WP 1.4
 \item start date: 05.07.2016
 \item end date: 05.01.2017
\end{itemize}

\subsubsection{Develop a microcontroller for the Wavy}
\begin{itemize}
 \item goal(s): A microcontroller to handle and identify brain waves.
 \item description: Develop a microcontroller thats manage the detection and identification of brain waves. It is working with the developed algorithm and sending it over a bluetooth interface to a chosen device. The communication with the controller is handled with the developed software interface.
 \item expected results: A working microcontroller.
 \item responsible person: Hardware engineer
 \item dependencies: WP 2.2, WP 2.3
 \item start date: 05.01.2017
 \item end date: 05.07.2017
\end{itemize}

\subsection{Prototype development}
\begin{itemize}
 \item goal(s): A device that is similar than the device for the consumer.
 \item description: Develop a prototype which can satisfy all requirements
 \item expected results:
 \begin{itemize}
  \item Wavy prototype
  \item Demonstration software
  \item A study confirming the superiority of our new method
 \end{itemize}
 \item responsible person: Hardware engineer
 \item dependencies: WP 2
 \item start date: 05.07.2017
 \item end date: 05.07.2018
\end{itemize}

\subsubsection{Create the Wavy prototype}
\begin{itemize}
 \item goal(s): Develop a wearable prototype.
 \item description: The Wavy prototype doesn’t have to be the final product but it should function like specified above.
 \item expected results: working prototype (Wavy)
 \item responsible person: Hardware engineer
 \item dependencies: WP 2
 \item start date: 05.07.2017
 \item end date: 05.03.2018
\end{itemize}

\subsubsection{Write a software prototype}
\begin{itemize}
 \item goal(s): Write a simple prototype including all essential features of WaveMeIn.
 \item description: The software prototype should implement the software, which is required for using the WaveMeIn Features.
 \item expected results: A small software to show how the Wavy device works (log in to an operation system).
 \item responsible person: Software engineer
 \item dependencies: WP 2
 \item start date: 05.07.2017
 \item end date: 05.03.2018
\end{itemize}

\subsubsection{Comparing with existing authentication methods}
\begin{itemize}
 \item goal(s): certificate, publicity
 \item description: An external research comparison.
 \item expected results: Confirm that the safety of this authentication method is superior to any other authentication method.
 \item responsible person: Computer science researcher
 \item dependencies: WP 2, WP 3.1
 \item start date: 05.03.2018
 \item end date: 05.07.2018
\end{itemize}