### 2015-01-08 proposal/sections/costs.tex
Costs

\section{Costs}
\label{sect:costs}
\subsection{Person}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Person & Commitment & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Project Manager & 100\% & 54000 & • & • & • & • \\ 
\hline 
Quality Manager & 75\% & 27000 & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
\end{tabular}

\subsection{Hardware}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Commercial EEG Headsets & 3000 & • & • & • & • \\ 
\hline 
Medical EEG Device & 60000 & • & • & • & • \\ 
\hline 
Server & 4000 & • & • & • & • \\ 
\hline 
Workstations & Notebooks & 6400 & • & • & • & • \\ 
\hline 
\end{tabular}
\subsection{Material}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Hardware for the Prototype & 2000 & • & • & • & 2000 \\ 
\hline 
EEG Sensors & 500 & • & • & • & 500 \\ 
\hline 
• & • & • & • & • & • \\ 
\hline 
\end{tabular} 
\subsection{Travel Costs}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Conferences & 4000 & • & 4000 & • & 8000 \\ 
\hline 
EEG Sensors & 500 & • & • & • & 500 \\ 
\hline 
• & • & • & • & • & • \\ 
\hline 
\end{tabular} 
\subsection{Other Costs}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Teambuilding, Snacks, Coffee, Meetings & 7000 & 7000 & 7000 & 3500 & 8000 \\
\hline 
\end{tabular}