### 2015-01-08 proposal/sections/implication-risk.tex
Costs

\section{Expected Implications and Risks}
\label{sect:implication-risk}

\subsection*{Implications and challenges}
One of the biggest challenge in the project is the basic research to detect unique brain waves. When our research has success and we are able to detect it, there are many use cases in other topics. One use case could be usage for high secure applications and systems. Where biometrical methods has some disadvantages, we have discussed above, our method can reduce this disadvantages.
Other topics could be the medicine. Current we know only a little bit about brain waves. With our research we could increase our medical understanding of brain waves.

\subsection*{Risks}

\subsubsection*{Less money}
\begin{itemize}
 \item Description: Get less money to realize the project than defined in the cost accounting.
 \item Solution: Our project is calculated tightly. So we must revalidate the hole project and create a new proposal with less features and work packages.
\end{itemize}

\subsubsection*{More money needed}
\begin{itemize}
 \item Description: The project needs more money than defined in the cost accounting.
 \item Solution: Looking for new investors to get more money. Sell know-how to other companies or research groups.
\end{itemize}

\subsubsection*{Less time for research}
\begin{itemize}
 \item Description: The brain waves research needs more time than supposed.
 \item Solution: Buy know-how from other companies and research groups. Get external experts to increase the research.
\end{itemize}

\subsubsection*{Team members leave}
\begin{itemize}
 \item Description: One or more team members leave the project.
 \item Solution: All work topics have two members. To avoid loosing knowledge, there are meetings every week to beware it. After knowledge about leaving a team member, we are looking for a new team member.
\end{itemize}

\subsubsection*{Brain waves location}
\begin{itemize}
 \item  Description: It is not possible to detect brain waves in the near of the ear.
 \item Solution: We are using sensors to detect brain waves at the forehead.
\end{itemize}

\subsubsection*{Brain waves detection}
\begin{itemize}
 \item Description: It is not possible to detect uniqueness of brain waves.
 \item Solution: It is not possible to continue, because this is the base of the project.
\end{itemize}

\subsubsection*{Misinterpretation}
\begin{itemize}
 \item Description: The team members not understand the specification correctly.
 \item Solution: Make meetings and reviews to ensure that every team member has understand it correctly.
\end{itemize}