### 2015-01-08 proposal/sections/costs.tex
Costs

\section{Costs}
\label{sect:costs}
\subsection{Personnel}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline 
Person & \% & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Project Manager & 100\% & 54000,00 & 55080,00 & 56181,60 & 28652,62 & 193914,22 \\ 
\hline 
Quality Manager & 75\% & 27000,00 & 27540,00 & 28090,80 & 14326,21 & 96957,11 \\ 
\hline 
Neuroscience Researcher 1 & 100\% & 41810,00 & 42646,20 & - & - & 84856,20 \\ 
\hline 
Neuroscience Researcher 2 & 100\% & 41810,00 & 42646,20 & - & - & 84856,20 \\ 
\hline 
Computer Science R. 1 & 100\% & 41810,00 & 42646,20 & 43499,12 & 22184,55 & 150139,88 \\ 
\hline 
Computer Science R. 2 & 100\% & 41810,00 & 42646,20 & 43499,12 & 22184,55 & 150139,88 \\ 
\hline 
Hardware Engineer 1 & 100\% & 31160,00 & 31783,20 & 32418,86 & 16533,62 & 111895,67 \\ 
\hline 
Hardware Engineer 2 & 100\% & 31160,00 & 31783,20 & 32418,86 & 16533,62 & 111895,67 \\ 
\hline 
Software Engineer 1 & 100\% & 31160,00 & 31783,20 & - & - & 62943,20 \\ 
\hline 
Software Engineer 2 & 100\% & 31160,00 & 31783,20 & - & - & 62943,20 \\ 
\hline 
\textbf{SUM} & & 372880,00 & 380337,60 & 236108,38 & 120415,27 & 989325,98 \\ 
\hline 
\end{tabular}

\subsection{Hardware}
\begin{tabular}{|l|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Commercial EEG Headsets & 3000,00 & - & - & - & 3000,00 \\ 
\hline 
Medical EEG Device & 60000,00 & - & - & - & 60000,00 \\ 
\hline 
Server & 4000 & - & - & - & 4000,00 \\ 
\hline 
Workstations/Notebooks & 6400,00 & - & - & - & 6400,00 \\ 
\hline 
\textbf{SUM} & 77000,00 &  &  &  & 77000,00 \\ 
\hline 
\end{tabular}

\subsection{Material}
\begin{tabular}{|l|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Hardware for the Prototype & 2000,00 & - & - & - & 2000,00 \\ 
\hline 
EEG Sensors & 500,00 & - & - & - & 500,00 \\ 
\hline 
\textbf{SUM} & 2500,00 &  &  &  & 2500,00 \\ 
\hline 
\end{tabular}

\subsection{Travel Costs}
\begin{tabular}{|l|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Conferences & 4000,00 & - & 4000,00 & - & 8000,00 \\ 
\hline 
\textbf{SUM} & 4000,00 &  & 4000,00 &  & 8000,00 \\ 
\hline 
\end{tabular}

\subsection{Other Costs}
\begin{tabular}{|l|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Team-building, Snacks, Coffee, Meetings & 7000,00 & 7000,00 & 7000,00 & 3500,00 & 24500,00 \\
\hline 
\end{tabular}